### 2024-06-02 thesis/input_files/data_exploratory.tex
June 2nd 2024

\section{GEFCom2014}
The first class of datasets we identified are the ones from the GEFCom2014 competition. The data is freely hosted on Dr.\ Hong's blog \cite{hong2016probabilistic}.
The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
We provide forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consists of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the $1^{st}$ of January 2011 to the $15^{th}$ of June 2013, see Figure \ref{fig:price_track_fig1}.\ 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track GEFCom2014}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see Figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, Figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and Figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisations, we can see their goodness for predicting zonal prices since there exists a clear correlation between the candidate predictors and the target variables.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
    \caption{Zonal price versus zonal load GEFCom2014}
    \label{fig:gefcom_zonal_price_vs_zonal_load}
\end{figure}

@@

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while Figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, Figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

@@

Following, we use box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that weekdays and workdays involve a higher load and possess a wider range.

>>>>

\section{GEFCom2014}
The first class of datasets we identified are the ones from the GEFCom2014 competition. The data is freely hosted on Dr.\ Hong's blog \cite{hong2016probabilistic}.
The main reason is that, these datasets are considered an excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
We provide forecasts for the load and price quantities.
\\
\subsection{GEFCom2014 price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consists of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the $1^{st}$ of January 2011 to the $15^{th}$ of June 2013, see Figure \ref{fig:price_track_fig1}.\ 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The 15 target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013, 17/12/2013. 
The first three tasks were designed for testing, they did not count for the competition.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track GEFCom2014}
    \label{fig:price_track_fig1}
\end{figure}
\subsubsection{EDA}
In exploring this data we started off by plotting the time series of the zonal prices alongside with the total load and zonal load, see Figure \ref{fig:price_track_fig1}.
After that, we plotted the scatter plot for inspecting the correlation between the predictors and the dependent variables, Figure \ref{fig:gefcom_zonal_price_vs_zonal_load} and Figure \ref{fig:gefcom_zonal_price_vs_total_load}. From these visualisations, we can see their goodness for predicting zonal prices since there exists a clear correlation between the candidate predictors and the target variables.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/gefcom_zonal_price_vs_zonal_load.png}
    \caption{Zonal price versus zonal load of GEFCom2014}
    \label{fig:gefcom_zonal_price_vs_zonal_load}
\end{figure}

@@

\subsection{GEFCom2014 load track}
In the load track, contestants were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we have a txt file with the competition instructions and 15 subfolders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Any data from the various task folders are shifted by one month between each other.
Also in this track, the first three tasks do not count towards the ranking of the various algorithms.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_\-V2.csv file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
The pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.
\subsubsection{EDA}
We applied the same kind of EDA to the GEFCom load data. Figure \ref{fig:gefcom_load_historical} depicts the historical load while Figure \ref{fig:gefcom_w_avg_historical} reports the historical weather temperature. Finally, Figure \ref{fig:gefcom_load_vs_w_avg} shows the scatter plot between the load and the weather.
These plots motivate the selection of weather temperature as one of the predictors for the load.

@@

Following, we use box plots in order to visualize the relations between categorical independent variables and the dependent one. Such visualisation are reported in Figure \ref{fig:CH_day_of_week_boxplot_2021} and Figure \ref{fig:CH_is_holiday_boxplot_2021}. We can see that workdays involve a higher load and possess a wider range than weekdays and holidays.

### 2024-06-02 thesis/input_files/implementation.tex
June 2nd 2024

\subsection{K-nearest neighbours}
The object KNeighborsRegressor of the \href{https://scikit-learn.org/stable/}{sklearn} module neighbors has been used.

@@

\subsection{Long short term memory}
The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library.

\subsection{Kernel ridge regression}
The object KernelRidge of the \href{https://scikit-learn.org/stable/}{sklearn} module kernel\_ridge has been used.

\subsection{Kernel support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used.
% by specifying rbf as the kernel parameter.

>>>>

\subsection{K-nearest neighbours}
The object KNeighborsRegressor of the \href{https://scikit-learn.org/stable/}{sklearn} module neighbors has been used with 12 neighbours and the Euclidean distance.

@@

\subsection{Long short term memory}
The LSTM predictor has been built using the \href{https://pytorch.org}{torch} library, see Section \ref{sec:lstm point} for architecture details.

\subsection{Kernel ridge regression}
The object KernelRidge of the \href{https://scikit-learn.org/stable/}{sklearn} module kernel\_ridge has been used with the Gaussian RBF kernel.

\subsection{Kernel support vector regression}
The object SVR of the \href{https://scikit-learn.org/stable/}{sklearn} module svm has been used with the Gaussian RBF kernel.
% by specifying rbf as the kernel parameter.

### 2024-06-02 thesis/input_files/literaturereview.tex
June 2nd 2024

In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term load forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF). Consensus in the literature is to use one day, two weeks and three years respectilvely \cite{hong_phd} as cut off horizons; see Figure \ref{fig:time} for a visualisation. 

@@

This section presents the results of the bibliometric analysis we performed on March 6th 2024. This survey has been carried out by using the Scopus citation database. For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.

@@

Hong et al. \cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting. It differentiates between techniques and methodologies. With techniques they refer to a family of models, like multiple linear regression or artificial neural networks. On the other hand, methodologies consist of general frameworks that can be incorporated into any method, for example variable selection mechanisms. Also this paper stresses the need for some guidelines to standardize research in the field.
Nowotarski et al. \cite{nowotarski} carried out a review of probabilistic forecasting.
Weron et al. \cite{lago} offer a set of best practices when forecasting electricity prices in order to have a common framework to evaluate and compare future research.
Zhang et al. \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
Ziel et al. \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 

@@

David et al. \cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance. Furthermore, they propose a recursive framework for parameter estimation.

@@

He et al. \cite{he2022cooperative} models  multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition (CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR). In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally kernel density estimation (KDE) converts quantile forecasts into density estimates.
Wan et al. \cite{wan2016direct} proposes a combination of QR and extreme learning machine (ELM) to generate non parametric probabilistic forecasts of wind generation.
Zhang et al. \cite{zhang2019wind} forecasts wind speed by adopting QRMGM, which is a combination of QR with a minimal gated memory network.
Hyndman et al. \cite{hyndman1996estimating} explains kernel estimator of conditional density, analyses its asymptotic behaviour and covers an application for the daily temperatures of Melbourne.
Kaur et al. \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
Van der Meer et al. \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.

@@

Marcjasz et al. \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the German market.
Nowotarski et al. \cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging. Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimized. The observed result is that QRA performed better compared to twelve individual models.
Arora et al. \cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional, hyndman1996estimating}. Their conclusion is that kernel density methods are competitive against exponential smoothing when forecasting residential data. Conversely, exponential smoothing has still an edge in predicting small to medium-sized enterprises data (SMEs).
Zhang et al. \cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting. The proposed methods behave well compared to an autoregressive ARMA model.

@@

Haben et al. \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting. Within this study, they focus on load forecasting at the low voltage level.

@@

Koochali et al. \cite{koochali2022random} reviews various existing methods for assessing probabilistic forecast models and discusses their advantages and disadvantages.
Matheson et al. \cite{matheson1976scoring} develops classes of scoring rules for continuous probability distributions.
Gneiting et al. \cite{gneiting2007strictly}

@@

Gneiting et al. \cite{gneiting2014probabilistic}

@@

Zhang et al \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting. They test it for different regression models such as random forest (RF), gradient boosting regression tree (GBRT), linear regression, and least squares support vector regression (LSSVM).
Jónsson et al. \cite{jonsson2014predictive}

@@

Fatema et al. \cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals estimation. Then, it inputs prediction intervals to kernel density estimation in order to estimate a probability distribution.

@@

Huurman et al. \cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the Danish market. Their empirical results suggest that weather is central for point forecasting day ahead prices. The opposite conclusion are drawn for density forecasting.

@@

The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al. \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.

@@

Xie et al. \cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA). 
Maciejowska et al. \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of quantile regression averaging (QRA) \cite{nowotarski2015computing}.
Haben et al. \cite{haben2016hybrid} mixes conditional kernel density (CKD) and quantile regression in their competition entry.
Gaillard et al. \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
Ziel et al. \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.

@@

\section{Kernel methods literature review}
%INTRO KERNEL METHODS E ENERGY PRICE MODELLING
Kernel methods are a class of algorithms for pattern analysis.
With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
Throughout this thesis, we will address the performance of kernel methods in the context of EF.
\\
%- Kernel history evolution.
Their name comes from the German word kern, which translates to core in english. Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term "definite kernels". Following, Hilbert's and Schmidt's work, \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
In 1938, Schoenberg \cite{schoenberg} developed the mathematical results that allow us to find the kernel associated to a specific feature space metric.
In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernels in linear spaces.
In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel Hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
It was in the 1990s that the theory of kernel methods got popular, particularly in the field of machine learning. Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
Nevertheless kernel methods received very little attention in the specific setting of EF literature.
\\
The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review} and \cite{shawe2004kernel}.
\\
Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}. Muandet et al. \cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings. 
It has to be said that, computing and storing such embeddings becomes prohibitive for large scale settings. Rudi et al. \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
Another example of the applicability of kernel methods is \cite{supersamples}. In that paper, Smola et al.\ kernelize the herding algorithm introduced in \cite{welling2009herding,welling2009herding,Welling2010}. The original herding algorithm converts empirical moments into a sequence of pseudo-samples approximating the probability density function (PDF) of the observed random variable. Its weakness is that it can only take a finite number of features and this limits its expressive power.
Smola et al.\ overcome this by translating the algorithm into the framework of kernels.

>>>>

In terms of forecasting horizons, we can group EF into four major categories: very short term forcasting (VSTF), short term forecasting (STF), medium term forecasting (MTF) and long term forecasting (LTF). Consensus in the literature is to use one day, two weeks and three years respectilvely as cut off horizons \cite{hong_phd}; see Figure \ref{fig:time} for a visualisation. 

@@

This section presents the results of the bibliometric analysis we performed on March $6^th$ 2024. This survey has been carried out by using the Scopus citation database. For details on the specific queries entered in Scopus, refer to Appendix \ref{src_code}.

@@

Hong et al.\ \cite{prob_elf} discusses the state of the art in probabilistic electric load forecasting. It differentiates between techniques and methodologies. With techniques they refer to a family of models, like multiple linear regression or artificial neural networks. On the other hand, methodologies consist of general frameworks that can be incorporated into any method, for example variable selection mechanisms. Also this paper stresses the need for some guidelines to standardize research in the field.
Nowotarski et al.\ \cite{nowotarski} carried out a review of probabilistic forecasting.
Weron et al.\ \cite{lago} offer a set of best practices when forecasting electricity prices in order to have a common framework to evaluate and compare future research.
Zhang et al.\ \cite{zhang2014review} considers state of the art methods in wind power probabilistic forecasting and describes current challenges and possible future developments.
Ziel et al.\ \cite{ziel2018probabilistic} provides detailed tables, groups research papers by the time dimension and objective and reports dataset, model and accuracy 

@@

David et al.\ \cite{david2016probabilistic} adopt a combination of autoregressive moving average (ARMA) and generalized autoregressive conditional heteroskedasticity (GARCH) in probabilistic forecasts of solar irradiance. Furthermore, they propose a recursive framework for parameter estimation.

@@

He et al.\ \cite{he2022cooperative} models multistep wind speed probabilistic forecasting by mixing complementary ensemble empirical mode decomposition (CEEDMAN), least absolute shrinkage and selection operator (LASSO) and quantile regression (QR). In this work, CEEMDAN is used to decompose the wind speed time series, LASSO compresses high dimensional features, QR is used for obtaining quantile forecasts, finally kernel density estimation (KDE) converts quantile forecasts into density estimates.
Wan et al.\ \cite{wan2016direct} proposes a combination of QR and extreme learning machine (ELM) to generate non parametric probabilistic forecasts of wind generation.
Zhang et al.\ \cite{zhang2019wind} forecasts wind speed by adopting QRMGM, which is a combination of QR with a minimal gated memory network.
Hyndman et al.\ \cite{hyndman1996estimating} explains kernel estimator of conditional density, analyses its asymptotic behaviour and covers an application for the daily temperatures of Melbourne.
Kaur et al.\ \cite{kaur2022energy} carries out a comparative study of techniques spanning ideas from statistic and artificial intelligence.
Van der Meer et al.\ \cite{van2018review} provides another thorough anlysis of the probabilistic forecasting realm by covering recent advances and identifying research gaps.

@@

Marcjasz et al.\ \cite{probablistic_electricity_forecast2} uses distributional neural networks to create probabilistic forecasts for the day ahead electricity prices in the German market.
Nowotarski et al.\ \cite{nowotarski2015computing} introduce a method for constructing prediction intervals (PI) and call it quantile regression averaging. Their idea is weighting a set of model predictions such that the pinball loss of the weighted model is minimized. The observed result is that QRA performed better compared to twelve individual models.
Arora et al.\ \cite{arora2016forecasting} focus on modelling electricity smart meter data by proposing a non parametric probabilistic technique based on kernel density estimation and conditional kernel density estimation \cite{rosenblatt1969conditional, hyndman1996estimating}. Their conclusion is that kernel density methods are competitive against exponential smoothing when forecasting residential data. Conversely, exponential smoothing has still an edge in predicting small to medium-sized enterprises data (SMEs).
Zhang et al.\ \cite{zhang2020probability} introduce a framework based on quantile regression and kernel density estimation in the context of short term wind forecasting. The proposed methods behave well compared to an autoregressive ARMA model.

@@

Haben et al.\ \cite{haben2018short} analyses a variety of techniques in terms of both probabilistic and point forecasting. Within this study, they focus on load forecasting at the low voltage level.

@@

Koochali et al.\ \cite{koochali2022random} reviews various existing methods for assessing probabilistic forecast models and discusses their advantages and disadvantages.
Matheson et al.\ \cite{matheson1976scoring} develops classes of scoring rules for continuous probability distributions.
Gneiting et al.\ \cite{gneiting2007strictly}

@@

Gneiting et al.\ \cite{gneiting2014probabilistic}

@@

Zhang et al.\ \cite{zhang2020two} proposes a two stage bootstrap sampling framework for probabilistic load forecasting. They test it for different regression models such as random forest (RF), gradient boosting regression tree (GBRT), linear regression, and least squares support vector regression (LSSVM).
Jónsson et al.\ \cite{jonsson2014predictive}

@@

Fatema et al.\ \cite{fatema2023probabilistic} considers Gaussian process regression for point forecasting and prediction intervals estimation. Then, it inputs prediction intervals to kernel density estimation in order to estimate a probability distribution.

@@

Huurman et al.\ \cite{huurman2012power} surveys the predictive power of weather variables for electricity prices in the Danish market. Their empirical results suggest that weather is central for point forecasting day ahead prices. The opposite conclusion are drawn for density forecasting.

@@

The focus of GEFCom 2014 was on probabilistic forecasting, Hong et al.\ \cite{hong2016probabilistic} discusses the problem tracks, the data and the winning methods.

@@

Xie et al.\ \cite{xie2016gefcom2014} propose a two stage approach; in the first stage they use multiple linear regression (MLR) to build a point forecast, then in the second stage they try different approaches for modelling the MLR residuals, among other they tried exponential smoothing (ESM), artificial neural networks (ANN) and autoregressive integrated moving average (ARIMA). 
Maciejowska et al.\ \cite{maciejowska2016hybrid} proposes a new probabilistic model extending the idea of quantile regression averaging (QRA) \cite{nowotarski2015computing}.
Haben et al.\ \cite{haben2016hybrid} mixes conditional kernel density (CKD) and quantile regression in their competition entry.
Gaillard et al.\ \cite{gaillard2016additive,gaillardasemi} combines quantile regression with generalized additive models \cite{hastie2017generalized}.
Ziel et al.\ \cite{ziel2016lasso} estimates an AR model through the LASSO \cite{tibshirani1996regression} instead of the standard OLS.

@@

\section{Kernel methods literature review}
%INTRO KERNEL METHODS E ENERGY PRICE MODELLING
Kernel methods are a class of algorithms for pattern analysis.
With kernel methods we are able to apply linear methods with predictors in a high dimensional space, without having to explicitly evaluate the involved dot products of the features.
Throughout this thesis, we will address the performance of kernel methods in the context of EF.
\\
%- Kernel history evolution.
Their name comes from the German word kern, which translates to core in english. Such term was first used by David Hilbert in his paper on integral equations \cite{hilbert} where he introduced the term "definite kernels". Following, Hilbert's and Schmidt's work, \cite{schmidt} lead to the introduction of a new space, the Hilbert space.
In 1909, James Mercer improved Hilbert's work by proposing his theorem \cite{mercer}. This theorem underlies the power of kernel methods, that is the kernel trick.
In 1938, Schoenberg \cite{schoenberg} developed the mathematical results that allow us to find the kernel associated to a specific feature space metric.
In 1941, Kolmogorov \cite{kolmogorov} carried out stuides on representing kernels in linear spaces.
In 1950, Aronszajn \cite{aronszajn} published the first work on reproducing kernel Hilbert spaces (RKHS); developing the general method for representing kernels in linear spaces.
In 1964, Aizerman \cite{aizerman} further improved the theory of RKHS.
It was in the 1990s that the theory of kernel methods got popular, particularly in the field of machine learning. Kernels have been used in various different tasks such as support vector machines (SVM) \cite{vapnik1} \cite{vapnik2}, Gaussian process classifiers \cite{williams}, spline methods \cite{wahba}, neural networks \cite{poggio} and principal component analysis \cite{pca_scholkopf}.
Nevertheless kernel methods received very little attention in the specific setting of EF literature.
\\
The kernel theory needed for this thesis work is covered in Section \ref{kernel_theory}. 
For an introduction to kernel methods, we refer to \cite{learning_with_kernels, hofmann2006review, shawe2004kernel}.
\\
Kanagawa et Fukumizu introduces to the concept of kernel mean embedding \cite{pmlr}. Muandet et al.\ \cite{Muandet_2017} surveys established results and new advances in the theory of Hilbert space distribution embeddings. 
It has to be said that, computing and storing such embeddings becomes prohibitive for large scale settings. Rudi et al.\ \cite{2022nystrom} proposes an efficient approximation procedure based on the Nyström method \cite{nystrom}, providing also an upper bound for the approximation error.
Another example of the applicability of kernel methods is \cite{supersamples}. In that paper, Smola et al.\ kernelize the herding algorithm introduced in \cite{welling2009herding,welling2009herding,Welling2010}. The original herding algorithm converts empirical moments into a sequence of pseudo-samples approximating the probability density function (PDF) of the observed random variable. Its weakness is that it can only take a finite number of features and this limits its expressive power.
Smola et al.\ overcome this by translating the algorithm into the framework of kernels.
